Rank Verification for Exponential Families

### 2016-04-22 verifying-winner.tex
Minor fixes

A Schur-concave function is symmetric by default since $a \succeq b$ and $b \succeq a$ if and only if $b$ is a permutation of the coordinates of $a$. Conversely a symmetric and log-concave function is Schur-concave \cite{Marshall:2010hb}, thus many common exponential family distributions lie in this class. Schur-concave functions indeed appear a lot, \Cref{sec:examples} includes both continuous and discrete examples.

@@

\begin{proof}[Proof of \Cref{thm:seltest}]
Without loss of generality, otherwise by relabeling, we consider only the case where the winner is $1$ and the runner-up is $2$, etc., i.e.\ $X_1 \ge \cdots \ge X_n$. In this case, we will test the null hypothesis $H_{01}: \theta_1 \le \max_{j > 1} \theta_j$, which is the union of the null hypotheses $H_{01j}: \theta_1 \le \theta_j$ for $j > 1$. For each of these we can construct an exact $p$-value $p_{1j}$. Hence by \Cref{lma:union}, a test based on $p_{1*} = \max_j p_{1j}$ would be exact for $H_{01}$. Procedure 1 will perform an unadjusted pairwise test comparing $X_1$ to $X_2$. Hence it is sufficient to show that $p_{12} = p_{1*}$ and that the test based on $p_{12}$ coincides with the unadjusted pairwise test.

@@

The proof above holds true if the exponential family is defined on the counting measure. If ties are broken independently and randomly, as in \Cref{sec:multinomial_example}, the end points on the rays can be considered as ``half an atom''.

@@

Gupta and Nagel \cite{Gupta:1967wg} devised a rule to select a subset that includes the maximum $\pi_j$. In other words, if the selected subset is $J\left(X\right)$, they guaranteed
$$\PP\left(\argmax_j \pi_j \in J\left(X\right)\right) \ge 1 - \alpha.$$
This was achieved by finding an integer $D$, as a function on $m$, $n$ and $\alpha$, and select by
$$J\left(X\right) = \left\{j: X_j \ge \max_k X_k - D\right\}.$$
Gupta and Nagel also provided the method to choose $D$. The subset selection can be modified. Inherently, the event $\left\{J\left(X\right) = \left\{1\right\}\right\}$ is the same as inferring $\pi_1$ is the largest. We thus implemented their test for this purpose.

\Cref{fig:power} gives the power curves for $\text{Multinomial}\left(m, \pi\right)$ and
$$\pi \propto \left(e^\delta, 1, \ldots, 1\right),$$
for various combinations of $m$ and $n$. For Gupta and Nagel, we use $\alpha = 0.05$; but in light of the extra factor of $\frac{n-1}{n}$ in \Cref{eqn:marginal}, we will apply the selective procedure with $\frac{n}{n-1} \alpha$ such that the marginal type I error rate of both procedures are controlled at $\alpha$. Gupta and Nagel coincides with our test at $n = 2$; however as $n$ grows, the selective test shoes significantly more power than Gupta and Nagel's test.

@@

\begin{proof}
Again we start with assuming $X_1 \ge \cdots \ge X_n$. The $p$-values in question are derived from the conditional law
$$\mathcal{L}_{\theta_1 - \theta_j = \delta} \left(D_{1j} \middle| D_{1j}, X_2, \ldots, X_n, A\right),$$
which is the truncated distribution
\begin{align*}
D_{1j} & \sim \exp\left(\left(\theta_1 - \theta_j\right) D_{1j} + \theta_2 X_2 + \cdots + \left(\theta_1 + \theta_j\right) M_{1j} + \cdots + \theta_n X_n \right) \\
& ~~~~~~~~ g\left(M_{ij} + D_{ij}, X_2, \ldots, M_{ij} - D_{ij}, \ldots X_n\right) 1_{A_1} \\
& \sim \exp\left(\delta D_{ij}\right) g\left(M_{ij} + D_{ij}, X_2, \ldots, M_{ij} - D_{ij}, \ldots X_n\right) 1_{A_1}.
\end{align*}

@@

\begin{proof}
If the confidence bound $\delta^*$ is negative, we report $-\infty$ so it definitely is valid and conservative. On the other hand, if $\delta^* \ne 0$, then at $\delta^*$ the $p$-value $p_{12} = \alpha$. The denominator is the positive part of a positively skewed distribution, and thus bigger than half of the distribution. So the unadjusted pairwise test $p$-value from Procedure 2 comparing $X_{[1]}$ and $X_{[2]}$ is larger than $p_{12}$ from Procedure 2', rendering Procedure 2 conservative.
\end{proof}

@@

This can be fitted into the selective sequential model selection framework proposed by Fithian, Taylor and Tibshirani \cite{Fithian:2015uj}. Given the observations $X_1 \ge \cdots \ge X_n$, drawn from an unknown distribution $F$, we consider the sequence of null hypothesis of
$$H_{0j}: \theta_j \le \max_{k > j} \theta_k.$$
Alternatively, we can think of them as a sequence of nested model
$$M_0\left(X\right) \subseteq M_1\left(X\right) \subseteq \cdots \subseteq M_n\left(X\right), ~~~~~~~~ M_j\left(X\right)^c = \left\{\theta_1 > \cdots > \theta_j > \max_{k > j} \theta_k\right\},$$
and the each $M_j$ depends on $X$ through the order of $X$. The step-down procedure for rejecting $H_{0j}$ is equivalent to selecting the smallest $j$ for which $F \in M_j\left(X\right)$.

@@

Hence similar to the proof of \Cref{thm:seltest}, Schur-concavity ensures $p_{j\left(j+1\right)} \ge p_{jk}$ for all $k>j$, meaning that it is sufficient to compare $X_j$ and $X_{j+1}$. In fact, we can take $p_j = p_{j\left(j+1\right)}$. These are valid selective $p$-value as required in Fithian, Taylor and Tibshirani, as
$$\PP_F \left(p_j \le \alpha \middle| M_{j-1}\left(X\right)\right) \le \alpha \text{ a.s.}, ~~~~~~~~ F \in M_{j-1}\left(X\right)$$
by tower property.

The conditioning above indeed holds some significance. The sufficient filtration \cite{Fithian:2015uj} $\mathscr{F}_k$ is given as the $\sigma$-field
\begin{align*}
\mathscr{F}_k & = \sigma\left(M_0\left(X\right), \ldots, M_j\left(X\right), X_1, \ldots, X_{j-1}, \frac{X_j + X_{j+1}}{2}, X_{j+2}, \ldots, X_n\right) \\
& = \sigma\left(\left\{X_1 > \ldots > X_j > \max_{k>j} X_k\right\}, X_1, \ldots, X_{j-1}, \frac{X_j + X_{j+1}}{2}, X_{j+2}, \ldots, X_n\right) \\
& = \sigma\left(M_j\left(X\right), X_1, \ldots, X_{j-1}, \frac{X_j + X_{j+1}}{2}, X_{j+2}, \ldots, X_n\right),
\end{align*}
which is the conditioning we used just now. Hence the nested sequence of model satisfies the subpath sufficiency principle, and the $p$-values sequence $p_j$ is independent on nulls, as defined in and shown in Theorem 4 of \cite{Fithian:2015uj}. Thus the step-down procedure where we keep rejecting $H_{0j}$ until the first time that $p_j > \alpha$ controls FWER.
\end{proof}

@@

\begin{proof}
The $p$-values $p_{j\left(j+1\right)}$ obtained in Procedure 3' are always smaller than their counterpart in Procedure 3, as the upper truncation at $X_{j-1}$ is on the upper tail. Therefore Procedure 3 is more conservative and definitely valid.
\end{proof}

>>>>

A Schur-concave function is symmetric by default since $a \succeq b$ and $b \succeq a$ if and only if $b$ is a permutation of the coordinates of $a$. Conversely a symmetric and log-concave function is Schur-concave \citep{Marshall:2010hb}, thus many common exponential family distributions lie in this class. Schur-concave functions indeed appear a lot, \Cref{sec:examples} includes both continuous and discrete examples.

@@

\begin{proof}
Without loss of generality, otherwise by relabeling, we consider only the case where the winner is $1$ and the runner-up is $2$, etc., i.e.\ $X_1 \ge \cdots \ge X_n$. In this case, we will test the null hypothesis $H_{01}: \theta_1 \le \max_{j > 1} \theta_j$, which is the union of the null hypotheses $H_{01j}: \theta_1 \le \theta_j$ for $j > 1$. For each of these we can construct an exact $p$-value $p_{1j}$. Hence by \Cref{lma:union}, a test based on $p_{1*} = \max_j p_{1j}$ would be exact for $H_{01}$. Procedure 1 will perform an unadjusted pairwise test comparing $X_1$ to $X_2$. Hence it is sufficient to show that $p_{12} = p_{1*}$ and that the test based on $p_{12}$ coincides with the unadjusted pairwise test.

@@

The proof above holds true if the exponential family is defined on the counting measure. If ties are broken independently and randomly, the end points on the rays can be considered as ``half an atom''.

@@

\citet{Gupta:1967wg} devised a rule to select a subset that includes the maximum $\pi_j$. In other words, if the selected subset is $J\left(X\right)$, they guaranteed
$$\PP\left(\argmax_j \pi_j \in J\left(X\right)\right) \ge 1 - \alpha.$$
This was achieved by finding an integer $D$, as a function on $m$, $n$ and $\alpha$, and select by
$$J\left(X\right) = \left\{j: X_j \ge \max_k X_k - D\right\}.$$
They also provided the method to choose $D$. The subset selection can be modified. Inherently, the event $\left\{J\left(X\right) = \left\{1\right\}\right\}$ is the same as inferring $\pi_1$ is the largest. We thus implemented their test for this purpose.

\Cref{fig:power} gives the power curves for $\text{Multinomial}\left(m, \pi\right)$ and
$$\pi \propto \left(e^\delta, 1, \ldots, 1\right),$$
for various combinations of $m$ and $n$. For their method, we use $\alpha = 0.05$; but in light of the extra factor of $\frac{n-1}{n}$ in \Cref{eqn:marginal}, we will apply the selective procedure with $\frac{n}{n-1} \alpha$ such that the marginal type I error rate of both procedures are controlled at $\alpha$. Their test coincides with our test at $n = 2$; however as $n$ grows, the selective test shows significantly more power than Gupta and Nagel's test.

@@

\begin{proof}
Again we start with assuming $X_1 \ge \cdots \ge X_n$. The $p$-values in question are derived from the conditional law
$$\mathcal{L}_{\theta_1 - \theta_j = \delta} \left(D_{1j} \middle| D_{1j}, X_2, \ldots, X_n, A\right),$$
which is the truncated distribution
\begin{align*}
D_{1j} & \sim \exp\left(\left(\theta_1 - \theta_j\right) D_{1j} + \theta_2 X_2 + \cdots + \left(\theta_1 + \theta_j\right) M_{1j} + \cdots + \theta_n X_n \right) \\
& ~~~~~~~~ g\left(M_{1j} + D_{1j}, X_2, \ldots, M_{1j} - D_{1j}, \ldots X_n\right) 1_{A_1} \\
& \sim \exp\left(\delta D_{1j}\right) g\left(M_{1j} + D_{1j}, X_2, \ldots, M_{1j} - D_{1j}, \ldots X_n\right) 1_{A_1}.
\end{align*}

@@

\begin{proof}
If the confidence bound $\delta^*$ is negative, we report $-\infty$ so it definitely is valid and conservative. On the other hand, if $\delta^* \ne 0$, then at $\delta^*$ the $p$-value $p_{12} = \alpha$. The denominator is the positive part of a positively skewed distribution, and thus bigger than half of the distribution. So the unadjusted pairwise test $p$-value from Procedure 2 comparing $X_1$ and $X_2$ is larger than $p_{12}$ from Procedure 2', rendering Procedure 2 conservative.
\end{proof}

@@

This can be fitted into the selective sequential model selection framework proposed by \citet{Fithian:2015uj}. Given the observations $X_1 \ge \cdots \ge X_n$, drawn from an unknown distribution $F$, we consider the sequence of null hypothesis of
$$H_{0j}: \theta_j \le \max_{k > j} \theta_k.$$
Alternatively, we can think of them as a sequence of nested model
$$M_0\left(X\right) \subseteq M_1\left(X\right) \subseteq \cdots \subseteq M_n\left(X\right), ~~~~~~~~ M_j\left(X\right)^c = \left\{\theta_1 > \cdots > \theta_j > \max_{k > j} \theta_k\right\},$$
and the each $M_j$ depends on $X$ through the order of $X$. The step-down procedure for rejecting $H_{0j}$ is equivalent to selecting the smallest $j$ for which $F \in M_j\left(X\right)$.

@@

Hence similar to the proof of \Cref{thm:seltest}, Schur-concavity ensures $p_{j\left(j+1\right)} \ge p_{jk}$ for all $k>j$, meaning that it is sufficient to compare $X_j$ and $X_{j+1}$. In fact, we can take $p_j = p_{j\left(j+1\right)}$. These are valid selective $p$-value as required in \citet{Fithian:2015uj}, as
$$\PP_F \left(p_j \le \alpha \middle| M_{j-1}\left(X\right)\right) \le \alpha \text{ a.s.}, ~~~~~~~~ F \in M_{j-1}\left(X\right)$$
by tower property.

The conditioning above indeed holds some significance. The sufficient filtration \citep{Fithian:2015uj} $\mathscr{F}_k$ is given as the $\sigma$-field
\begin{align*}
\mathscr{F}_k & = \sigma\left(M_0\left(X\right), \ldots, M_j\left(X\right), X_1, \ldots, X_{j-1}, \frac{X_j + X_{j+1}}{2}, X_{j+2}, \ldots, X_n\right) \\
& = \sigma\left(\left\{X_1 > \ldots > X_j > \max_{k>j} X_k\right\}, X_1, \ldots, X_{j-1}, \frac{X_j + X_{j+1}}{2}, X_{j+2}, \ldots, X_n\right) \\
& = \sigma\left(M_j\left(X\right), X_1, \ldots, X_{j-1}, \frac{X_j + X_{j+1}}{2}, X_{j+2}, \ldots, X_n\right),
\end{align*}
which is the conditioning we used just now. Hence the nested sequence of model satisfies the subpath sufficiency principle, and the $p$-values sequence $p_j$ is independent on nulls, as defined in and shown in Theorem 4 of \citet{Fithian:2015uj}. Thus the step-down procedure where we keep rejecting $H_{0j}$ until the first time that $p_j > \alpha$ controls FWER.
\end{proof}

@@

\begin{proof}
The $p$-values $p_{j\left(j+1\right)}$ obtained in Procedure 3' are always smaller than their counterpart in Procedure 3, as the upper truncation at $X_{j-1}$ is on the upper tail. Therefore Procedure 3 is conservative and definitely valid.
\end{proof}
